---
title: "Building multivariate visualizations with ggplot2"
output: 
  html_document: 
    fig_height: 3.5
---

```{r message=FALSE}
library(tidyverse)
library(scales)
```

### Warm up: Which of the two datasets does this visualization use? Determine which variable is mapped to which aesthetic (x-axis, y-axis, etc.) element of the dataset.

```{r echo=FALSE, fig.height=2}
daily <- read_csv("data/daily.csv")

ggplot(data = daily, aes(x = ride_date, y = n_rides, color = day_of_week)) +
  geom_smooth(se = FALSE, method = "loess") +
  theme_minimal() +
  labs(x = "Ride date", y = "Number of rides", color = "Day of week",
       title = "Daily rides", subtitle = "Durham, NC") +
  scale_color_viridis_d()
```

### Exercise 1: Using information from https://ggplot2.tidyverse.org/articles/ggplot2-specs.html add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment. Do different things happen when you map aesthetics to  discrete and continuous variables? What happens when you use more than one aesthetic?

```{r}
ggplot(data = daily, aes(x = ride_date, y = n_rides, color = day_of_week)) +
  geom_smooth(method = "loess", se = FALSE) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(x = "Ride date", y = "Number of rides", color = "Day of week",
       title = "Daily rides", subtitle = "Durham, NC")
```

### Exercise 2: What is wrong with the following?

```{r eval=FALSE}
daily %>%
  mutate(day_type = if_else(day_of_week %in% c("Sat", "Sun"), "Weekend", "Weekday")) %>%
  ggplot(aes(x = ride_date, y = n_rides, color = day_type)) %>%
    geom_point()
```

### Exercise 3: Work with your neighbor to sketch what the following plot will look like. No cheating! Do not run the code, just think through the code for the time being.

```{r eval=FALSE}
low_weekdays <- daily %>%
  filter(!(day_of_week %in% c("Sat", "Sun")) & n_rides < 100)

ggplot(daily, aes(x = ride_date, y = n_rides, color = day_of_week)) +
  geom_point(data = low_weekdays, size = 5, color = "gray") +
  geom_point() +
  geom_text(data = low_weekdays, aes(y = n_rides + 15, label = ride_date), 
            size = 2, color = "black")
```

### Exercise 4: How would you fix the following plot to color by day of week?

```{r}
ggplot(daily, aes(x = ride_date, y = n_rides, color = day_of_week)) +
  geom_smooth(color = "blue")
```

### Exercise 5: In what ways is the following plot different than the previous?

```{r}
durham_voters <- read_csv("data/durham_voters.csv")

durham_voters_summary <- durham_voters %>%
  group_by(race_code, gender_code, age) %>%
  summarize(n_all_voters = n(), 
            n_rep_voters = sum(party == "REP")) %>%
  filter(gender_code %in% c("F", "M") & 
         race_code %in% c("W", "B", "A") &
         age != "Age < 18 Or Invalid Birth Date")

durham_voters_summary %>% 
  pivot_longer(cols = starts_with("n_"), 
               names_to = "voter_type", values_to = "n", 
               names_prefix = "n_") %>%
  mutate(age_cat = as.numeric(as.factor(age))) %>%
  ggplot(aes(x = age, y = n, color = voter_type)) + 
  geom_point() + 
  geom_line(aes(x = age_cat)) +
  facet_grid(race_code ~ gender_code, scales = "free_y") +
  expand_limits(y = 0) +
  theme(legend.position = "bottom")
```

### Exercise 6: Fix the axis labels in the following plot so they don't overlap by playing around with their orientation.

```{r fig.width=4}
ggplot(durham_voters_summary, aes(x = age, y = n_all_voters)) +
  geom_bar(stat = "identity") +
  facet_grid(race_code ~ gender_code, scales = "free_y")
```
